Body Mass Index and Pathway Dysregulation in Cancer

### 2016-09-12 presentation/beamer_theme/example.tex
draft ctcr presentation

% \usepackage{pgfpages}
% \pgfpagesuselayout{2 on 1}[a4paper,border shrink=3mm]

@@

\title[BMI and Cancer]{Body Mass Index and Pathway Dysregulation in Cancer}
\subtitle{CTCR Talk}
\author[Riku Takei]{Riku Takei (MSc)\\Supervisor: Assoc. Prof. Mik Black}
\institute{Department of Biochemistry, University of Otago}

@@

@@

\begin{frame}
	\frametitle{Body Mass Index and Cancer}
	\begin{columns}
		\column{0.6\textwidth}
		{\footnotesize
			Cancers are caused by dysregulation in various pathways/mechanisms that allow them to survive and proliferate
			\begin{itemize}
				\item Commonly know as the Hallmarks of Cancer
			\end{itemize}
			Many reports on obesity being a major risk for cancer and poor prognosis
		}
		\column{0.5\textwidth}
		\includegraphics[width=0.45\paperwidth]{./images/cancerhallmark}
	\end{columns}
\end{frame}

@@

\begin{frame}
	\frametitle{Obesity-associated Genetic Signatures}
	Creighton \textit{et al.} (2012)
	\begin{itemize}
		\item Identified obesity-associated genetic signatures from the tumours of breast cancer patients
		\item Showed that some compounds were able to specifically regulate the expression of these genetic signatures (potential for targeted therapy)
	\end{itemize}
\end{frame}

\begin{frame}{{\normalsize Obesity-associated genes from Creighton \textit{et al.} (2012) }}
	103 samples used in the study, 799 obesity-associated gene probes (695 unique genes) identified\\
	{\footnotesize
		\begin{itemize}
			\item Majority of these genes were downregulated in obese patients
		\end{itemize}
	}
	\begin{center}
		\includegraphics[width=0.6\paperwidth]{./images/creighton}
	\end{center}
\end{frame}

\begin{frame}{Obesity-associated genes from Fuentes-Mattei \textit{et al.}}
	\begin{columns}
		\column{0.5\textwidth}
		Fuentes-Mattei \textit{et al.} (2014)
		\begin{itemize}
			\item 137 ER+ breast cancer patients in the study, 130 obesity-associated genes identified
			\item Direct evidence of how obesity/adiposity contributes to tumour growth (using mouse model)
		\end{itemize}
		\column{0.5\textwidth}
		\includegraphics[width=0.45\paperwidth]{./images/fmcircos}
	\end{columns}
\end{frame}

\begin{frame}
	\frametitle{Methods}
	Downloaded RNA-seq data from 8 different cancer types from ICGC
	\vskip2ex
	Microarray data from Creighton's paper
	\vskip2ex
	Microarray data from Fuentes-Mattei's paper
	\vskip2ex
	Microarray data from Cris Print
	\vskip2ex
	\includegraphics[width=0.4\textwidth]{./images/tcga}
	\hfill
	\includegraphics[width=0.4\textwidth]{./images/icgc}
\end{frame}

\begin{frame}{Validation of Methodology -- Creighton \textit{et al.} (2012)}
	103 samples used in the study, 799 obesity-associated gene probes (695 unique genes) identified\\
	\vskip2ex
	Metagene was obtained from SVD using the genes that were common in both Creighton \textit{et al.} and other cancer data sets\\
	\begin{center}
		\includegraphics[width=0.6\paperwidth]{./images/creighton}
	\end{center}
\end{frame}

@@

\begin{frame}{Transformation Matrix}
	\begin{center}
		\includegraphics[height=0.7\paperheight,width=0.40\paperwidth]{./images/tmp/transmatexp1}
	\end{center}
\end{frame}

\begin{frame}{Transformation Matrix}
	\begin{center}
		\includegraphics[width=0.70\paperwidth]{./images/tmp/transmatexp2}
	\end{center}
\end{frame}

@@

\begin{frame}{{\normalsize Obesity-associated genes from Fuentes-Mattei \textit{et al.}}}
	\begin{columns}
		\column{0.5\textwidth}
		137 ER+ breast cancer samples in the study, 130 unique genes identified\\
		\vskip2ex
		Data were normalised  with RMA normalisation method\\
		\vskip2ex
		\column{0.5\textwidth}
		\includegraphics[width=0.45\paperwidth]{./images/fmcircos}
	\end{columns}
\end{frame}

@@

@@

\begin{frame}{Creighton \textit{et al.} data revisited}
	Gene expression analysis was carried out on the Creighton \textit{et al.} data:
	\begin{itemize}
		\item Original Creighton data
		\item Caucasian-only Creighton data
	\end{itemize}
	\vskip2ex
	Gene expression analysis was repeated with the residual version of the above data
	\vskip2ex
	Each ``obesity-associated genes'' were checked for any overlaps with the original Creighton metagene
\end{frame}

\begin{frame}{Completely remove the effect of ethnicity}
	African American patients are more likely to have higher BMI than Cauacasian patients
	\vskip2ex
	Removed the effect of ethnicity by removing the African American patients' data (total of 77 patient data to work with)
	{\footnotesize
		\begin{itemize}
			\item All of the other clinical variables were adjusted as before
		\end{itemize}
	}
\end{frame}

\begin{frame}{DEGs in residual data}
	\begin{center}
		\includegraphics[width=0.42\paperwidth]{./images/venn0}
		\includegraphics[width=0.45\paperwidth]{./images/venn1}
	\end{center}
\end{frame}

\begin{frame}{Metagene Analysis}
	All of the gene probe sets I have found (8 in total) were used to create metagenes as before
	\begin{itemize}
		\item Checked in Creighton \textit{et al.} data
		\item Transferred into other cancer data sets
	\end{itemize}
\end{frame}

@@

% \begin{frame}
% 	Repeated this with Cris' data, but same results as in Creighton and ICGC data
% \end{frame}

% \begin{frame}{Do the metagenes correlate with each other?}
% 	\begin{center}
% 		\includegraphics[width=0.45\paperwidth]{./images/cor1}
% 	\end{center}
% \end{frame}

% \begin{frame}{Common genes across multiple cancer types}
% 	DEGs between the obese and non-obese groups were identified for each cancer types
% 	\vskip2ex
% 	For each cancer types, samples were randomly split into two groups (same proportion as obese vs. non-obese samples)
% 	\vskip2ex
% 	{\footnotesize
% 		\begin{tabular}{ccccccccc}
% 			& \multicolumn{8}{c}{No. of cancer types the gene overlapped}\\
% 			\rule{0pt}{2.25ex}                    & 1    & 2    & 3   & 4  & 5 & 6 & 7 & 8 \\
% 			\hline
% 			\rule{0pt}{2.25ex}Obese vs. Non-obese & 6865 & 1907 & 289  & 26   & 1 & 0 & 0 & 0 \\
% 			Randomly split groups                 & 6202 & 1488 & 198  & 18   & 1 & 0 & 0 & 0 \\
% 			No. of genes found                    & 5012 & 870  & 89   & 6    & 0 & 0 & 0 & 0 \\
% 			on average                            &      &      &      &      &   &   &   & \\
% 			Empirical FDR                         & 0.73 & 0.46 & 0.31 & 0.23 & 0 & 0 & 0 & 0 \\
% 		\end{tabular}
% 	}
% \end{frame}

% \begin{frame}{Pathway Enrichment Analysis}
% 	Maybe each cancer type have a different set of genes activated
% 	\vskip2ex
% 	Are there any specific pathways enriched in any specific cancer type?
% 	\vskip2ex
% 	If so, is it common across other cancer types?
% 	\vskip2ex
% 	How relevant is the pathway to obesity and/or cancer?
% 	\vskip2ex
% 	Can we use it to narrow down our search for obesity-associated signature?
% \end{frame}

% \begin{frame}{Pathway Enrichment Analysis}
% 	Gene Ontology (GO) database was used to check for the function(s) being enriched by the DEGs in each cancer type (between obese and non-obese groups)
% 	\vskip2ex
% 	Used a rank-based approach and controlled the error rate using FDR
% 	\vskip2ex
% 	\begin{center}
% 		\textbf {No pathways showed up as significantly enriched}
% 	\end{center}
% \end{frame}

% \begin{frame}{Common genes and pathway enrichment analysis}
% 	There were no obesity-associated genes that were common across all the ICGC cancer types
% 	\vskip2ex
% 	There were no pathways enriched in any of the obesity-associated genes found so far
% \end{frame}

\begin{frame}{Relevance of BMI in Cris' data}
	\begin{center}
		\includegraphics[width=0.60\paperwidth]{./images/tmp/mikres}
	\end{center}
\end{frame}

\begin{frame}{So \ldots BMI is not related to cancer?}
	All the results so far suggests that there is no association between BMI and cancer
	\vskip2ex
	However, the metagenes are picking up gene signatures  that are common across all cancer types, but not related to BMI
	\vskip2ex
	So what signature/mechanism is the metagene picking up?
\end{frame}

\begin{frame}{Gatza \textit{et al.} (2010)}
	Used 18 pathways to define sub-class of breast tumours, based on the expression of the pathway signatures
	\begin{itemize}
		\item AKT, ER, PR, HER2, PI3K, p53, MYC, RAS, etc...
	\end{itemize}
	\vskip2ex
	Each pathway signature was obtained by overexpressing a single gene in an appropriate cell line, and then differentially expressed genes were noted down
	\begin{itemize}
			\item Use these pathways to see what the obesity-associated genes are most similar to
	\end{itemize}
\end{frame}

\begin{frame}{Tricky thing about metagene direction}
	Need to consider how the original data was manipulated:
	\begin{enumerate}
		\item Normalisation methods (RMA vs MAS5)
		\item Scaling the data or not
		\item Metagene scaling (probit vs rank-based)
	\end{enumerate}
	\vskip2ex
	Need to check the direction of each pathway metagenes with its gene (e.g. EGFR gene with EGFR pathway metagene)
	\vskip2ex
	Need to consider the grouping/clustering of the pathway metagenes
\end{frame}

\begin{frame}{What I mean about directions...}
	\includegraphics[width=0.45\paperwidth]{./images/output-0}
	\includegraphics[width=0.45\paperwidth]{./images/output-1}
\end{frame}

\begin{frame}{Some weird genes}
	\includegraphics[width=0.45\paperwidth]{./images/output-2}
	\includegraphics[width=0.45\paperwidth]{./images/output-3}
\end{frame}

\begin{frame}{Grouping/clustering of different pathways}
	\begin{center}
		\includegraphics[width=0.45\paperwidth]{./images/gatzares1}
	\end{center}
\end{frame}

\begin{frame}{Re-making results from Gatza \textit{et al.} paper}
	\includegraphics[width=0.45\paperwidth]{./images/out2}
	\includegraphics[width=0.40\paperwidth]{./images/gatzares1}
\end{frame}

\begin{frame}{Predicting obesity metagene with pathway metagene}
	Make a linear model (in Cris' data) with sample BMI and/or pathway metagenes and use this to predict Creighton's obesity metagene
	\vskip2ex
	\begin{itemize}
		\item If any of the pathway metagenes are significant in the model, the pathway is a good predictor of the obesity metagene
	\end{itemize}
	\vskip2ex
	Then use this model to predict the obesity metagene in Creighton data
	\begin{itemize}
		\item Compare the predicted obesity metagene with the SVD-generated obesity metagene
	\end{itemize}
\end{frame}

\begin{frame}{Results}
	Sample BMI was not significant  in the model
	\vskip2ex
	Only pathway that was a good predictor of obesity metagene was PR pathway metagene
\end{frame}

\begin{frame}{\normalsize Comparison of the predicted obesity metagene with true value}
	\begin{center}
		\includegraphics[width=0.45\paperwidth]{./images/tmp/prediction1}
		\includegraphics[width=0.45\paperwidth]{./images/tmp/prediction2}
	\end{center}
\end{frame}

\begin{frame}{\normalsize Comparison of the predicted obesity metagene with true value}
	\begin{center}
		\includegraphics[width=0.45\paperwidth]{./images/tmp/prediction3}
		\includegraphics[width=0.45\paperwidth]{./images/tmp/prediction4}
	\end{center}
\end{frame}

\begin{frame}{Overall Summary}
	\textbf{Creighton/FM metagene analysis}
	\begin{itemize}
		\item Both Creighton and FM metagenes correlated with the overall gene expression in any cancer data set
		\item The metagene score does not correlate with the BMI status of the samples
	\end{itemize}
	\vskip2ex
	\textbf{Gene expression and metagene analysis}
	\begin{itemize}
		\item Again, all of the gene sets found correlated with the overall gene expression in any data set, but none of them correlated with the BMI status
	\end{itemize}
\end{frame}

\begin{frame}{Overall Summary}
	\textbf{Gatza pathway metagene analysis}
	\begin{itemize}
		\item Sample BMI was not a good predictor of the obesity metagene, but PR pathway metagene was
		\item However, the model was not great at predicting the obesity metagene
			\begin{itemize}
				\item Other mechanism contributing to obesity metagene?
			\end{itemize}
	\end{itemize}
\end{frame}

% \begin{frame}{Discussion}
% 	Tumour heterogeneity within and between samples
% 	{\footnotesize
% 		\begin{itemize}
% 				% TODO rewrite the following sentence
% 			\item Not finding any significant genes or pathways because the samples are so different from one another
% 		\end{itemize}
% 	}
% 	\vskip2ex
% 	Tumour sample quality and size
% 	{\footnotesize
% 		\begin{itemize}
% 			\item Picking up too much noise and/or don't have enough samples to find anything significant (i.e. shit data)
% 		\end{itemize}
% 	}
% \end{frame}

% \begin{frame}{Discussion -- Inter-tumour heterogeneity 1}
% 	Each patient has an obesity driven pathway dysregulated, as well as different ``core'' pathways that lead to cancer progression
% 	\vskip2ex
% 	Each data set will have a different collection of patients, which in turn will have a different collection of dysregulated pathways
% 	{\footnotesize
% 		\begin{itemize}
% 			\item Hence DEGs show up, but not the pathways
% 		\end{itemize}
% 	}
% \end{frame}

% \begin{frame}{Discussion -- Inter-tumour heterogeneity 2}
% 	Some of the high BMI patients in the data set will have tumours that are BMI driven, but BMI may not be the driver for some of these patients
% 	\begin{itemize}
% 		\item Only subset of the patients with high BMI are actually driven by BMI
% 	\end{itemize}
% 	\vskip2ex
% 	Other data sets may have less of these truly BMI driven tumours
% 	\begin{itemize}
% 		\item Therefore no association with the signatures
% 	\end{itemize}
% \end{frame}

@@

% \begin{frame}
% 	\frametitle{References}
% 	\bibliographystyle{BiocRefStyle}
% 	\bibliography{/home/riku/Documents/References/BibTeX/MSc}
% \end{frame}

% \begin{frame}{test}
% 	\citet{Fuentes-Mattei2014}\\
% 	\citet{Creighton2012}\\
% 	\citet{Alexandrov2013}
% 	\cite{Gatza2010a}
% \end{frame}

>>>>

\usepackage{pgfpages}
\pgfpagesuselayout{2 on 1}[a4paper,border shrink=3mm]

@@

\title[BMI and Cancer Pathway]{Body Mass Index and Pathway Dysregulation in Cancer}
\subtitle{Committee Meeting 2016}
\author[Riku Takei]{Riku Takei (MSc)\\Supervisor: Assoc. Prof. Mik Black}
\institute{Department of Biochemistry, University of Otago}

@@

\begin{frame}{Overview}
	\begin{enumerate}
		\item Introduction to obesity and cancer
		\item Results from last committee meeting
		\item Results since the last committee meeting
		\item Future directions
	\end{enumerate}
\end{frame}

@@

\begin{frame}
	\frametitle{Body Mass Index and Cancer}
	\begin{columns}
		\column{0.6\textwidth}
		{\footnotesize
			Cancers are caused by dysregulation in various pathways/mechanisms that allow them to survive and proliferate
			\begin{itemize}
				\item Commonly know as the Hallmarks of Cancer \citep{Hanahan2011}
			\end{itemize}
			Many reports on obesity being a major risk for cancer and poor prognosis
		}
		\column{0.5\textwidth}
		\includegraphics[width=0.45\paperwidth]{./images/cancerhallmark}
	\end{columns}
\end{frame}

\begin{frame}
	\frametitle{Obesity-associated Genetic Signatures}
	Creighton \textit{et al.} (2012)
	\begin{itemize}
		\item Identified obesity-associated genetic signatures from the tumours of breast cancer patients
		\item Showed that some compounds were able to specifically regulate the expression of these genetic signatures (potential for targeted therapy)
	\end{itemize}
	Fuentes-Mattei \textit{et al.} (2014)
	\begin{itemize}
		\item Identified obesity-associated genetic signatures from the tumours of ER+ breast cancer patients
		\item Direct evidence of how obesity/adiposity contributes to tumour growth (using mouse model)
	\end{itemize}
\end{frame}

% \begin{frame}
% 	\frametitle{Genetic Signatures Across Multiple Cancer Types}
% 	\begin{center}
% 		\includegraphics[width=0.85\paperwidth]{./images/multicancersignature}
% 	\end{center}
% \end{frame}

@@

\begin{frame}{Progress Overview}
	Where I was at the last committee meeting:
	{\footnotesize
		\begin{itemize}
			\item Applied Creighton metagene to ICGC data
			\item Applied Fuentes-Mattei metagene to Creighton \textit{et al.} and ICGC data
			\item Found no association with BMI status
		\end{itemize}
	}
	\vskip2ex
	What I have done since the last committee meeting:
	{\footnotesize
		\begin{itemize}
			\item Gene expression analysis in Creighton \textit{et al.} data
			\item Gene expression analysis in ICGC data
			\item Pathway enrichment analysis in ICGC data
			\item Comparison of the obesity metagenes with other pathway metagenes from Gatza \textit{et al.} (preliminary results)
		\end{itemize}
	}
\end{frame}

\begin{frame}
	\frametitle{Methods}
	Download sequence (RNA-seq) data and clinical data from publicly available databases
	{\footnotesize
		\begin{itemize}
			\item The Cancer Genome Atlas (TCGA)
			\item International Cancer Genome Consortium (ICGC)
		\end{itemize}
	}
	\vskip2ex
	\includegraphics[width=0.4\textwidth]{./images/tcga}
	\hfill
	\includegraphics[width=0.4\textwidth]{./images/icgc}
\end{frame}

\begin{frame}{Clinical Data}
	Download all the clinical data available from TCGA databases
	\begin{itemize}
		\item {\footnotesize Clinical data available for 33 cancer types}\\
	\end{itemize}
	\vskip2ex
	Only interested in cancer types that have both height and weight data\\
	\begin{equation*}
		BMI = \frac{Weight (kg)}{Height^2 (m^2)}
	\end{equation*}
\end{frame}

\begin{frame}{Clinical Data with Patient BMI}
	Out of 33 cancer types, 14 types had both height and weight data\\
	{\footnotesize
		\begin{itemize}
			\item 8 cancer types had RNA-seq data\\
		\end{itemize}
	}
	\vskip2ex
	\begin{center}
		\begin{tabular}{lcccl}
			& Normal & Overweight & Obese & Total\\
			\rowcolor{LightCyan}
			BLCA & 103  & 97         & 61    & 261\\
			CESC & 85   & 62         & 77    & 224   \\
			\rowcolor{LightCyan}
			COAD & 75   & 80         & 71    & 226   \\
			KIRP & 26   & 53         & 45    & 124\\
			\rowcolor{LightCyan}
			LIHC & 140  & 75         & 49    & 264\\
			READ & 22   & 36         & 15    & 73\\
			\rowcolor{LightCyan}
			SKCM & 72   & 73         & 73    & 218\\
			UCEC & 90   & 106        & 286   & 482\\
			\hline
			\rule{0pt}{2.25ex}Total & 613 & 582 & 677 & 1872\\
		\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{Validation of Methodology -- Creighton \textit{et al.} (2012)}
	103 samples used in the study, 799 obesity-associated gene probes (695 unique genes) identified\\
	{\footnotesize
		\begin{itemize}
			\item Majority of these genes were downregulated in obese patients
		\end{itemize}
	}
	\begin{center}
		\includegraphics[width=0.6\paperwidth]{./images/creighton}
	\end{center}
\end{frame}

\begin{frame}{Validation of Methodology -- Creighton \textit{et al.} (2012)}
	Normalised the data using RMA normalisation method (\texttt{affy} package)\\
	\vskip2ex
	Metagene was obtained from SVD using the genes that were common in both Creighton \textit{et al.} and other cancer data sets\\
	\begin{center}
		\includegraphics[width=0.6\paperwidth]{./images/creighton}
	\end{center}
\end{frame}

@@

@@

\begin{frame}{Obesity-associated genes from Fuentes-Mattei \textit{et al.}}
	\begin{columns}
		\column{0.5\textwidth}
		137 ER+ breast cancer samples in the study, 130 unique genes identified\\
		\vskip2ex
		Data were normalised  with RMA normalisation method\\
		\vskip2ex
		Metagene was formed from the common genes between the Fuentes-Mattei \textit{et al.} and other cancer data sets (116 genes)
		\column{0.5\textwidth}
		\includegraphics[width=0.45\paperwidth]{./images/fmcircos}
	\end{columns}
\end{frame}

% \begin{frame}{Can the FM metagene predict BMI status?}
% 	If both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} found obesity-associated genetic signature, then these signatures should be similar
% 	{\footnotesize
% 		\begin{itemize}
% 			\item Therefore should be able to separate the samples based on BMI
% 		\end{itemize}
% 	}
% 	\vskip2ex
% 	No BMI information is available for Fuentes-Mattei \textit{et al.} data set, so you need to transform the Creighton \textit{et al.} with the metagene\\
% 	{\footnotesize
% 		\begin{itemize}
% 			\item Check whether the metagene correlates with BMI status
% 		\end{itemize}
% 	}
% \end{frame}

@@

\begin{frame}{Summary 1}
	Metagene from Creighton \textit{et al.} correlated with gene expression and patient BMI in Creighton data set
	\vskip2ex
	Metagene derived from Creighton's obesity-associated genes  correlated with the gene expression but not the patient BMI in ICGC data
	\vskip2ex
	Metagene derived from Fuentes-Mattei obesity-associated genes correlated with the gene expression but not the patient BMI in both Creighton \textit{et al.} and ICGC data
\end{frame}

@@

\begin{frame}{Creighton \textit{et al.} data revisited}
	Gene expression analysis was carried out on the Creighton \textit{et al.} data
	{\footnotesize
		\begin{itemize}
			\item in their paper, they found 799 gene probes with p \textless{} 0.01 and fold change \textgreater{} 1.2
		\end{itemize}
	}
	\vskip2ex
	1781 differentially expressed genes (p \textless{} 0.01) were found between obese vs. non-obese
	{\footnotesize
		\begin{itemize}
			\item very few genes had fold change \textgreater{} 1.2 (132 genes)
			\item no genes were found when p-value was adjusted
			\item top 799 genes were chosen from this list
		\end{itemize}
	}
	\vskip2ex
	239 gene probes were present in the top 799 gene probes that I have found
\end{frame}

\begin{frame}{Residual data}
	All of the non-BMI clinical variables were controlled for in the linear model
	{\footnotesize
		\begin{itemize}
			\item Age, ethnicity, menopause status, grade, LN status, ER/PR/HER2 statuses
		\end{itemize}
	}
	\vskip2ex
	Residual data (or remaining data after controlling for the variables) was used to find the DEGs
	\vskip2ex
	Again, more than 799 gene probes showed up (1104 genes)
	{\footnotesize
		\begin{itemize}
			\item 168 genes from Creighton \textit{et al.} were found in the top 799 genes
		\end{itemize}
	}
\end{frame}

\begin{frame}{DEGs in residual data}
	\begin{center}
		\includegraphics[width=0.50\paperwidth]{./images/venn0}
	\end{center}
\end{frame}

\begin{frame}{Completely remove the effect of ethnicity}
	African American patients are more likely to have higher BMI than Cauacasian patients
	\vskip2ex
	Removed the effect of ethnicity by removing the African American patients' data (total of 77 patient data to work with)
	{\footnotesize
		\begin{itemize}
			\item All of the other clinical variables were adjusted as before
		\end{itemize}
	}
	\vskip2ex
	Top 799 genes were identified in this data set (both raw and residual)
	{\footnotesize
		\begin{itemize}
			\item  148 and 92 genes were in the top 799 genes in the raw and residual data, respectively
		\end{itemize}
	}
\end{frame}

\begin{frame}{DEGs in Cauacasian-only data}
	\begin{center}
		\includegraphics[width=0.50\paperwidth]{./images/venn1}
	\end{center}
\end{frame}

\begin{frame}{Metagene Analysis}
	All of the gene probe sets I have found (8 in total) were used to create metagenes and to transform other cancer types
	\begin{itemize}
		\item Checked in Creighton \textit{et al.} data
		\item Transferred into other cancer data sets
	\end{itemize}
\end{frame}

@@

\begin{frame}{Do the metagenes correlate with each other?}
	\begin{center}
		\includegraphics[width=0.45\paperwidth]{./images/cor0}
		\includegraphics[width=0.45\paperwidth]{./images/cor1}
	\end{center}
\end{frame}

\begin{frame}{Metagenes are correlating with GE but not BMI}
	All of the metagenes found so far have correlated with the overall gene expression of the samples, but not with the sample BMI
	\vskip2ex
	The metagenes are picking up gene signatures  that are common across all cancer types, but not related to BMI
	\vskip2ex
	\begin{center}
		\textbf{So what signature/mechanism is the metagene picking up?}
	\end{center}
\end{frame}

\begin{frame}{Pathway signatures from Gatza \textit{et al.} (2010)}
	Made metagenes from the pathway signatures from the Gatza \textit{et al.} paper in ICGC data
	\vskip2ex
	Compared these pathway metagenes with the BMI metagenes
	{\footnotesize
		\begin{itemize}
			\item Is the BMI metagene similar to any of the pathway signatures?
		\end{itemize}
	}
	\vskip2ex
	All of the samples from the ICGC data were used (1872 samples in total)
\end{frame}

\begin{frame}{Results}
	\begin{center}
		\includegraphics[width=0.90\paperwidth]{./images/allmeta1}
	\end{center}
\end{frame}

\begin{frame}{Results}
	\begin{center}
		\includegraphics[width=0.90\paperwidth]{./images/allmeta2}
	\end{center}
\end{frame}

\begin{frame}{Results}
	\begin{center}
		\includegraphics[width=0.45\paperwidth]{./images/allmetacor1}
	\end{center}
\end{frame}

\begin{frame}{Summary 2}
	Metagenes found in Creighton \textit{et al.} data correlates with GE and BMI status in Creighton data, but not in  the ICGC data
	\vskip2ex
	The metagenes do not correlate with any other pathway metagenes
	\vskip2ex
	\begin{center}
		\textbf{Are there any common genes and/or pathways being dysregulated across all of the cancer types?}
	\end{center}
\end{frame}

\begin{frame}{Common genes across multiple cancer types}
	DEGs between the obese and non-obese groups were identified for each cancer types
	\vskip2ex
	For each cancer types, samples were randomly split into two groups (same proportion as obese vs. non-obese samples)
	\vskip2ex
	{\footnotesize
		\begin{tabular}{ccccccccc}
			& \multicolumn{8}{c}{No. of cancer types the gene overlapped}\\
			\rule{0pt}{2.25ex}                    & 1    & 2    & 3   & 4  & 5 & 6 & 7 & 8 \\
			\hline
			\rule{0pt}{2.25ex}Obese vs. Non-obese & 6865 & 1907 & 289  & 26   & 1 & 0 & 0 & 0 \\
			Randomly split groups                 & 6202 & 1488 & 198  & 18   & 1 & 0 & 0 & 0 \\
			No. of genes found                    & 5012 & 870  & 89   & 6    & 0 & 0 & 0 & 0 \\
			on average                            &      &      &      &      &   &   &   & \\
			Empirical FDR                         & 0.73 & 0.46 & 0.31 & 0.23 & 0 & 0 & 0 & 0 \\
		\end{tabular}
	}
\end{frame}

\begin{frame}{Pathway Enrichment Analysis}
	Maybe each cancer type have a different set of genes activated
	\vskip2ex
	Are there any specific pathways enriched in any specific cancer type?
	\vskip2ex
	If so, is it common across other cancer types?
	\vskip2ex
	How relevant is the pathway to obesity and/or cancer?
	\vskip2ex
	Can we use it to narrow down our search for obesity-associated signature?
\end{frame}

\begin{frame}{Pathway Enrichment Analysis}
	Gene Ontology (GO) database was used to check for the function(s) being enriched by the DEGs in each cancer type (between obese and non-obese groups)
	\vskip2ex
	Used a rank-based approach and controlled the error rate using FDR
	\vskip2ex
	\begin{center}
		\textbf {No pathways showed up as significantly enriched}
	\end{center}
\end{frame}

\begin{frame}{Summary}
	\textbf{Creighton/FM metagene analysis}
	\begin{itemize}
		\item Both Creighton and FM metagenes correlated with the overall gene expression in any cancer data set
		\item The metagene score does not correlate with the BMI status of the samples
	\end{itemize}
	\textbf{Gene expression and metagene analysis}
	\begin{itemize}
		\item Again, all of the gene sets found correlated with the overall gene expression in any data set, but none of them correlated with the BMI status
	\end{itemize}
	\textbf{Pathway enrichment analysis}
	\begin{itemize}
		\item There were no pathway significantly enriched in any cancer type
	\end{itemize}
\end{frame}

\begin{frame}{Discussion}
	Different methods used to find DEGs
	{\footnotesize
		\begin{itemize}
			\item Found more genes than Creighton \textit{et al.} because they were using different programs and normalisation method in their analyses
		\end{itemize}
	}
	\vskip2ex
	Tumour heterogeneity within and between samples
	{\footnotesize
		\begin{itemize}
			\item Not finding any significant genes or pathways because the samples are so different from one another
		\end{itemize}
	}
	\vskip2ex
	Tumour sample quality and size
	{\footnotesize
		\begin{itemize}
			\item Picking up too much noise and/or don't have enough samples to find anything significant
		\end{itemize}
	}
\end{frame}

\begin{frame}{Discussion -- Inter-tumour heterogeneity 1}
	Each patient has an obesity driven pathway dysregulated, as well as different ``core'' pathways that lead to cancer progression
	\vskip2ex
	Each data set will have a different collection of patients, which in turn will have a different collection of dysregulated pathways
	{\footnotesize
		\begin{itemize}
			\item Hence DEGs show up, but not the pathways
		\end{itemize}
	}
\end{frame}

\begin{frame}{Discussion -- Inter-tumour heterogeneity 2}
	Some of the high BMI patients in the data set will have tumours that are BMI driven, but BMI may not be the driver for some of these patients
	\begin{itemize}
		\item Only subset of the patients with high BMI are actually driven by BMI
	\end{itemize}
	\vskip2ex
	Other data sets may have less of these truly BMI driven tumours
	\begin{itemize}
		\item Therefore no association with the signatures
	\end{itemize}
\end{frame}

\begin{frame}{Conclusion}
	Obesity-associated genetic signatures cannot be used in other cancer types, nor in the same cancer type with different sample population
	\vskip2ex
	Gene expression and pathway enrichment analyses showed that there may not be any obesity-associated signature across multiple cancer types
\end{frame}

\begin{frame}{Future Directions}
	Explore the biology of the ``obesity-associated'' metagene
	{\footnotesize
		\begin{itemize}
			\item Use other pre-defined pathway signatures and see if any of them correlates with the metagene
		\end{itemize}
	}
	\vskip2ex
	Effect of BMI on methylation status in different cancer types
\end{frame}

@@

\begin{frame}
	\frametitle{References}
	\bibliographystyle{BiocRefStyle}
	\bibliography{/home/riku/Documents/References/BibTeX/MSc}
\end{frame}

\begin{frame}{test}
	\citet{Fuentes-Mattei2014}\\
	\citet{Creighton2012}\\
	\citet{Alexandrov2013}
	\cite{Gatza2010a}
\end{frame}
